---
title: "Qpc-GWASpanel.Rmd"
author: "em"
date: "July 9, 2018"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(viridis)
library(qpctools)
library(qvalue)
setwd('~/Documents/qpc-maize/') #sorry world

```

This notebook has code for running a form of Q<sub>pc</sub> analysis in a GWAS panel of 240 maize lines. 

### 1. read in the kinship matrix and trait data

```{r readinandprocessdata}
#read in the kinship matrix
myK = read.table('data/All_240E.nomaf.nomissing.K')
myKnames = read.table('data/240.names', stringsAsFactors = F)$V1
row.names(myK) = myKnames[1:dim(myK)[1]]
heatmap(as.matrix(myK), col=viridis(256))

#do the eigen decomposition
eigF = eigen(myK)
myU = eigF$vectors
myLambdas = eigF$values

#read in trait values
bluptable= read.table('data/blups_240', header=T)
mytraitsnice = sapply(names(bluptable)[-1], function(x){
  strsplit(x, '0607')[[1]][1]
}) ##neaten up the trait names
names(bluptable) = c('line',mytraitsnice)
```

### 2. calculate the PC cutoffs we're using for tests of selection

```{r pccutoff}
varexp = myLambdas/sum(myLambdas)
sumexp = sapply(1:length(varexp), function(x){sum(varexp[1:x])})

#get cutoffs for how many pcs to look at
pcmax = which(sumexp > 0.3)[1]



plot(myLambdas, bty="n", xlab = "PCs", ylab = "Eigenvalues")
abline(v = pcmax, col = viridis(6)[3], lwd=2)

tailCutoff=0
```

### 3. We can test for excess variation along specific PCs.
```{bash}
cat qpctools/R/calcQpc.R
```

```{r onepctest}
mydfs = apply(bluptable[,-1], 2, function(x){calcQpc(
  myZ = x, myU = eigF$vectors, myLambdas = eigF$values, myPCcutoff=0.3, vapcs = 119, tailCutoff = 1
)})

```


### 4. Make plots and look at the data

```{r summarizeresults}

getqvalues <- function(ptable){
qobj = qvalue(p = c(ptable))
myqvals = matrix(qobj$qvalues, nrow=dim(ptable)[1])
return(myqvals)
}

allpvals = sapply(1:22, function(x){mydfs[[x]]$pvals})
myqvals = getqvalues(allpvals)

#mysig = apply(allqvals, 2, function(x){ cut(x, c(0,0.001,0.01,0.05,0.1,1), labels=F)})
layout(matrix(1, nrow=1, ncol=1))
mysig2 =  cut((1:1000/1000), c(0,0.001,0.01,0.05,0.1,1)) #for legend
par(mar=c(8,14,2,2), xpd=TRUE, mfrow=c(1,1))
mycol = c(viridis(6, direction=1)[1:4], "white")
image(myqvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=pcmax), las=2, label=1:pcmax)
axis(2, at=(0:21)/21, labels = mytraitsnice, las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)
```

### 7. More plots looking at correlations between PCs and traits.

```{r pccor}
#function to calculate confidence intervals. This is mainly useful for plotting and these do not go into the actual statistical test.

calcCIs <- function(myName, myBlups=bluptable, myU=eigF$vectors, myLambdas=eigF$values){
  myZ = myBlups[,myName][1:239]
  myZ = myZ - mean(myZ)
  myBm = myZ %*% myU
  myCm = myBm/sqrt(myLambdas)
  myVa = var0(myCm[(119):238])
  myCI = sqrt(myVa*myLambdas)
  return(myCI)}

par(mar=c(5,5,1,1), mfrow=c(1,1))

#read in population data and merge with traits and eigenvectors
fgtable = read.table('data/FlintGarciaTableS1.csv', sep=',', header=T, stringsAsFactors = F)
mydf = data.frame(bluptable[-nrow(bluptable),], eigF$vectors, stringsAsFactors = F)
mydf$fgname = sapply(as.character(mydf$line), function(x){strsplit(x,'_')[[1]][2]})
fgmerge = dplyr::inner_join(mydf, fgtable, by=c('fgname'='Inbred'))

palette(viridis(6))
nicepops = c('mixed','non-stiff stalk','popcorn','stiff-stalk','sweet','tropical')
indvar = sqrt(diag(as.matrix(myK)))

myCIsDTS = calcCIs('DaysToSilk')
myCIsTKN = calcCIs('TotalKernelNo')


save(mydfs, allpvals,myqvals, mytraitsnice, fgmerge, myCIsDTS, myCIsTKN,file="data/qpc-maize_results.rda")

par(xpd=FALSE)
plot(fgmerge$X1, fgmerge$DaysToSilk, bty="n", xlab = "PC 1", ylab = "Days to Female Flowering", yaxt="n", col=as.factor(fgmerge$Subpopulation), lwd=2)
axis(2, las=2)
legend('topleft', nicepops, bty="n", pch=1, pt.lwd=2, col = viridis(6))
abline(lm(fgmerge$DaysToSilk ~ fgmerge$X1), col=viridis(6)[1], lwd=2)
abline(a=mean(fgmerge$DaysToSilk), b = 1.96*myCIsDTS[1], lty=2, col=viridis(6)[3], lwd=2)
abline(a=mean(fgmerge$DaysToSilk), b = -1.96*myCIsDTS[1], lty=2, col=viridis(6)[3], lwd=2)

plot(fgmerge$X10, fgmerge$DaysToSilk, bty="n", xlab = "PC 10", ylab = "Days to Female Flowering", yaxt="n", col=as.factor(fgmerge$Subpopulation), lwd=2)
axis(2, las=2)
legend('topright', nicepops, bty="n", pch=1, pt.lwd=2, col = viridis(6))
abline(lm(fgmerge$DaysToSilk ~ fgmerge$X10), col=viridis(6)[1], lwd=2)
abline(a=mean(fgmerge$DaysToSilk), b = 1.96*myCIsDTS[10], lty=2, col=viridis(6)[3], lwd=2)
abline(a=mean(fgmerge$DaysToSilk), b = -1.96*myCIsDTS[10], lty=2, col=viridis(6)[3], lwd=2)


plot(fgmerge$X3, fgmerge$DaysToSilk, bty="n", xlab = "PC 3", ylab = "Days to Female Flowering", yaxt="n", col=as.factor(fgmerge$Subpopulation), lwd=2)
axis(2, las=2)
legend('topleft', nicepops, bty="n", pch=1, pt.lwd=2, col = viridis(6))
abline(lm(fgmerge$DaysToSilk ~ fgmerge$X3), col=viridis(6)[1], lwd=2)
abline(a=mean(fgmerge$DaysToSilk), b = 1.96*myCIsDTS[3], lty=2, col=viridis(6)[3], lwd=2)
abline(a=mean(fgmerge$DaysToSilk), b = -1.96*myCIsDTS[3], lty=2, col=viridis(6)[3], lwd=2)


```


Checking the sensitivity to choice of PCs.


```{r}
## try with all PCS below the cutoff 
mydfsAllpcs = apply(bluptable[,-1], 2, function(x){calcQpc(
  myZ = x, myU = eigF$vectors, myLambdas = eigF$values, myPCcutoff=0.3, vapcs = 217, tailCutoff = 1
)})

allpvalsAllpcs = sapply(1:22, function(x){mydfsAllpcs[[x]]$pvals})
myqvalsAllpcs = getqvalues(allpvalsAllpcs)

#use only 50 PCs
mydfs50pcs = apply(bluptable[,-1], 2, function(x){calcQpc(
  myZ = x, myU = eigF$vectors, myLambdas = eigF$values, myPCcutoff=0.3, vapcs = 50, tailCutoff = 1
)})

allpvals50pcs = sapply(1:22, function(x){mydfs50pcs[[x]]$pvals})
myqvals50pcs = getqvalues(allpvals50pcs)


layout(matrix(1, nrow=1, ncol=1))
mysig2 =  cut((1:1000/1000), c(0,0.001,0.01,0.05,0.1,1)) #for legend
par(mar=c(8,10,2,2), xpd=TRUE, mfrow=c(1,1), mfrow=c(1,3))
mycol = c(viridis(6, direction=1)[1:4], "white")

## changing the number of PCs used
image(myqvalsAllpcs, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=pcmax), las=2, label=1:pcmax)
axis(2, at=(0:21)/21, labels = mytraitsnice, las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

image(myqvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=pcmax), las=2, label=1:pcmax)
axis(2, at=(0:21)/21, labels = mytraitsnice, las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

image(myqvals50pcs, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=pcmax), las=2, label=1:pcmax)
axis(2, at=(0:21)/21, labels = mytraitsnice, las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

save(allpvals50pcs, allpvalsAllpcs, allpvals, myqvalsAllpcs, myqvals50pcs, myqvals, pcmax, mytraitsnice, file = "data/changingVaPCresults.rda")


```


