---
title: "Qpc-euro.Rmd"
author: "em"
date: "February 13, 2017"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(viridis)
library(qpctools)
library(qvalue)
setwd('~/Documents/qpc-maize/') #sorry world
```

This notebook has code for running a form of Q<sub>pc</sub> analysis using the Major-Goodman GWAS panel and detecting selection in the European landraces.

# Load in the kinship matrix and make a conditional matrix 
```{r kmatrix}
load('data/euro.282.E.rda')
sigma11 = as.matrix(myF[1:906,1:906])
sigma12 = as.matrix(myF[1:906,907:1168])
sigma21 = as.matrix(myF[907:1168,1:906])
sigma22 = as.matrix(myF[907:1168,907:1168]) #we are dropping the last row
sigma.cond = sigma11 - sigma12 %*% solve(sigma22) %*% sigma21
condEig = eigen(sigma.cond)
cEigVectors = condEig$vectors
cEigValues = condEig$values
#save(cEigVectors, cEigValues, file = 'data/euro.282.condeig.rda')

traitNames = read.table('data/blup.names', stringsAsFactors=F)$V1

niceTraitnames = sapply(traitNames, function(x){
  strsplit(x, '0607')[[1]][1]
})

```

# Calculate cutoffs for PCs
```{r pccutoffs}
#get cutoffs for pcs based on % variation explained
varexp = cEigValues/sum(cEigValues)
sumexp = sapply(1:length(varexp), function(x){sum(varexp[1:x])})



#get cutoffs for how many pcs to look at
pcmax = which(sumexp > 0.30)[1]
pcmax

library(nFactors)
myscree = nScree(cEigValues)
myscree$Components[1]

library('AssocTests')
mytw = tw(cEigValues, eigenL = length(cEigValues))
mytw$SigntEigenL



plot(cEigValues, bty="n", xlab = "PCs", ylab = "Eigenvalues")
abline(v = pcmax, col = viridis(6)[3], lwd=2)
abline(v = myscree$Components[1], col = viridis(6)[4], lwd=2)
abline(v = mytw$SigntEigenL, col = viridis(6)[5], lwd=2)

legend('topright',c('30% var explained','Castells rule','Tracy Widom'), col = viridis(6)[3:5], lwd=2, bty="n")


```

# Run Qpc
The function is in qpctools/R/QpcEuro.R
```{bash}
cat qpctools/R/QpcEuro.R
```


```{r runqpc, eval=F}
qpceuroOut = lapply(1:22,function(x){Qpceuro(myI = x, vapcs = 906/2, tailCutoff = 1)})
save(qpceuroOut, file = "data/qpc_euro_output.rda")

```

# Look at results
```{r heatmap-figure}
load('data/qpc_euro_output.rda')

pcpvals = sapply(qpceuroOut, function(x) {x$pprime}) #matrix, rows are pvals, columns are traits
allqvals = get_q_values(pcpvals)

layout(matrix(1, nrow=1, ncol=1))
mysig2 =  cut((1:1000/1000), c(0,0.001,0.01,0.05,0.1,1)) #for legend
par(mar=c(8,14,2,2), xpd=TRUE, mfrow=c(1,1))
mycol = c(viridis(6, direction=1)[1:4], "white")
image(allqvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=nrow(pcpvals)), label=1:nrow(pcpvals))
axis(2, at=(0:21)/21, labels = niceTraitnames, las=2)
legend(-0.2,-0.15, c('FDR',levels(mysig2)), fill=c('white',mycol), border=c('white',rep('black',5)), bty="n", horiz=T)

image(pcpvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=nrow(pcpvals)), label=1:nrow(pcpvals))
axis(2, at=(0:21)/21, labels = niceTraitnames, las=2)
legend(-0.2,-0.15, c('P value',levels(mysig2)), fill=c('white',mycol), border=c('white',rep('black',5)), bty="n", horiz=T)

```


# Running the original version of the test.
The function is in qpctools/R/Qpceuro-nocond.R
```{bash}
cat qpctools/R/QpcEuro-nocond.R
```


```{r no-conditional-test, eval=F}
#
load('data/euroOnlyK.rda')

### function for running Qpc on European polygenic scores without the conditional test
ncEuroOut = lapply(1:22,function(x){Qpceuro_nocond(myI = x, vapcs = 906/2, tailCutoff = 1)})
save(ncEuroOut, file="data/qpc-euro-nc.rda")

```

```{r}
load('data/qpc-euro-nc.rda')

### look at the output
ncpvals = sapply(ncEuroOut, function(x) {x$pprime}) #matrix, rows are pvals, columns are traits
ncqvals = get_q_values(ncpvals)

#mysig = apply(allqvals, 2, function(x){ cut(x, c(0,0.001,0.01,0.05,0.1,1), labels=F)})
mysig2 =  cut((1:1000/1000), c(0,0.001,0.01,0.05,0.1,1)) #for legend
par(mar=c(8,14,2,2), xpd=TRUE, mfrow=c(1,1))
mycol = c(viridis(6, direction=1)[1:4], "white")
image(ncqvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=nrow(ncpvals)), label=1:nrow(ncpvals))
axis(2, at=(0:21)/21, labels = niceTraitnames, las=2)
legend(-0.2,-0.15, c('FDR',levels(mysig2)), fill=c('white',mycol), border=c('white',rep('black',5)), bty="n", horiz=T)

image(ncpvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=nrow(ncpvals)), label=1:nrow(ncpvals))
axis(2, at=(0:21)/21, labels = niceTraitnames, las=2)
legend(-0.2,-0.15, c('P value',levels(mysig2)), fill=c('white',mycol), border=c('white',rep('black',5)), bty="n", horiz=T)


```


```{r latdata, eval=F, include=F}
par(mar=c(6,6,3,3), mfrow=c(1,1))

#read in data about the landraces
eurodat = read.table('data/eurolandraceinfo.csv', sep=',', head=T, stringsAsFactors=F)
euronames = read.table('data/merged263Landraces.names', stringsAsFactors=F)[1:906,]


#get the conditional expectations for polygenic values
zconds = sapply(qpceuroOut, function(x){x$muprime})
colnames(zconds) = niceTraitnames
breedingvals = sapply(qpceuroOut, function(x){x$bv})
colnames(breedingvals) = niceTraitnames

zdf = data.frame(names=euronames, breedingvals-zconds, stringsAsFactors = F)
zdf$Code = sapply(zdf$names, function(x){substr(x, 1,2)})
mymerge = dplyr::left_join(zdf, eurodat, by="Code")

save(allqvals, mymerge, niceTraitnames, sigma11, file="data/euro_qpc_data.rda")

```

# Map of the landraces
```{r}
#lat long plot
library(maps)
load("data/euro_qpc_data.rda")
eurodat = read.table('data/eurolandraceinfo.csv', sep=',', head=T, stringsAsFactors=F)

map("world", xlim=c(1.5*min(eurodat$Longitude),1.2*max(eurodat$Longitude)), ylim=c(0.85*min(eurodat$Latitude),1.05*max(eurodat$Latitude)), fill=T, col="azure")
points(eurodat$Longitude, eurodat$Latitude, col = magma(6)[4], lwd=4)


```



