---
title: "Qpc-euro.Rmd"
author: "em"
date: "February 13, 2017"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(viridis)
library(qpctools)
library(qvalue)
setwd('~/Documents/qpc/') #sorry world
```

This notebook has code for running a form of Q<sub>pc</sub> analysis using the Major-Goodman GWAS panel and detecting selection in the European landraces.

First, load in the kinship matrix and make a conditional matrix 
```{r kmatrix}
load('data/euro.282.E.rda')


sigma11 = as.matrix(myF[1:906,1:906])
sigma12 = as.matrix(myF[1:906,907:1168])
sigma21 = as.matrix(myF[907:1168,1:906])
sigma22 = as.matrix(myF[907:1168,907:1168]) #we are dropping the last row
sigma.cond = sigma11 - sigma12 %*% solve(sigma22) %*% sigma21
condEig = eigen(sigma.cond)
cEigVectors = condEig$vectors
cEigValues = condEig$values
#save(cEigVectors, cEigValues, file = 'data/euro.282.condeig.rda')

traitNames = read.table('data/blup.names', stringsAsFactors=F)$V1

niceTraitnames = sapply(traitNames, function(x){
  strsplit(x, '0607')[[1]][1]
})

```


```{r make-eigens, eval=F}
#make eigendecomp files and save for later
#euroOnlyF = as.matrix(read.table('data/mergedLandraces.gmriE.5e4.K'))
#euroOnlyeigen = eigen(euroOnlyF)
#save(euroOnlyF, euroOnlyeigen, file='data/euroOnlyK.rda')

euro282eigen = eigen(myF)
save(euro282eigen, file = 'data/euro.282.eig.rda')
```


```{r pccutoffs}
#get cutoffs for pcs based on % variation explained
varexp = cEigValues/sum(cEigValues)
sumexp = sapply(1:length(varexp), function(x){sum(varexp[1:x])})
par(mfrow=c(1,2), mar=c(5,5,1,1))

#get cutoffs for how many pcs to look at
pcmax = which(sumexp > 0.30)[1]

#look at the distributions of eigenvals and percent var explained
par(mfrow=c(1,2), mar=c(5,5,1,1))
plot(varexp, bty="n", xlab = "PC Index", ylab = "variation explained by each eigenvector")
plot(sumexp,bty="n", xlab = "PC Index", ylab = "cumulative variation explained by efirst X eigenvectors")
abline(v=c(pcmax, tailCutoff), col = viridis(6)[2:4])

```

This is how we calculate Qpc
```{r}
Qpceuro

```


Running Qpc on traits
```{r runqpc, eval=F}
qpceuroOut = lapply(1:22,Qpceuro)
save(qpceuroOut, file = "data/qpc_euro_output.rda")
```


```{r heatmap-figure}
load('data/qpc_euro_output.rda')

pcpvals = sapply(qpceuroOut, function(x) {x$pprime}) #matrix, rows are pvals, columns are traits
allqvals = get_q_values(pcpvals)

layout(matrix(1, nrow=1, ncol=1))
mysig2 =  cut((1:1000/1000), c(0,0.001,0.01,0.05,0.1,1)) #for legend
par(mar=c(8,14,2,2), xpd=TRUE, mfrow=c(1,1))
mycol = c(viridis(6, direction=1)[1:4], "white")
image(allqvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=nrow(pcpvals)), label=1:nrow(pcpvals))
axis(2, at=(0:21)/21, labels = niceTraitnames, las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

image(pcpvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=nrow(pcpvals)), label=1:nrow(pcpvals))
axis(2, at=(0:21)/21, labels = niceTraitnames, las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)
```

```{r pccor-figure, eval=F}
par(mar=c(6,6,3,3), mfrow=c(1,1))
eurodat = read.table('data/eurolandraceinfo.csv', sep=',', head=T, stringsAsFactors=F)
euronames = read.table('data/merged263Landraces.names', stringsAsFactors=F)[1:906,]


zconds = sapply(qxpceuroOut, function(x){x$muprime})
colnames(zconds) = niceTraitnames
breedingvals = sapply(qxpceuroOut, function(x){x$bv})
colnames(breedingvals) = niceTraitnames

zdf = data.frame(names=euronames, breedingvals-zconds, stringsAsFactors = F)
zdf$Code = sapply(zdf$names, function(x){substr(x, 1,2)})
mymerge = dplyr::left_join(zdf, eurodat, by="Code")


save(allqvals, mymerge, niceTraitnames, sigma11, file="data/euro_qxpc_data.rda")

palette(viridis(3))

myI=22
plot(cEigVectors[,1], mymerge[,niceTraitnames[myI]]/diag(sigma11), bty="n", xlab = "Conditional PC 1", ylab = "Difference btw breeding value and expectation", col=as.factor(mymerge$Type), lwd=2, main=niceTraitnames[myI])
legend("bottomright", levels(as.factor(mymerge$Type)), col = viridis(3)[1:2], pch=1, pt.lwd=2, bty="n", cex=1.5)

palette(viridis(75))
plot(cEigVectors[,1], mymerge[,niceTraitnames[myI]]/diag(sigma11), bty="n", xlab = "Conditional PC 1", ylab = "Difference btw breeding value and expectation", col=as.factor(mymerge$Latitude), lwd=2, main=niceTraitnames[myI])
#legend('bottomright', as.character(round(c(quantile(mymerge$Latitude, probs=seq(0,1,1/3), na.rm=T)))), pch=1, bty='n', col=viridis(75)[round(c(quantile(mymerge$Latitude, probs=seq(0,1,1/3), na.rm=T)))], pt.lwd=2)

plot(mymerge$Latitude, mymerge[,niceTraitnames[myI]]/diag(sigma11), bty="n", xlab = "Latitude", ylab = "Difference btw breeding value and expectation", col=as.factor(mymerge$Type), lwd=2, main=niceTraitnames[myI])
abline(lm(mymerge[,niceTraitnames[myI]]~ mymerge$Latitude), lwd=2)

myI=1
palette(viridis(3))
plot(cEigVectors[,15], mymerge[,niceTraitnames[myI]]/diag(sigma11), bty="n", xlab = "Conditional PC 15", ylab = "Difference btw breeding value and expectation", col=as.factor(mymerge$Type), lwd=2, main=niceTraitnames[myI])
legend("bottomleft", levels(as.factor(mymerge$Type)), col = viridis(3)[1:2], pch=1, pt.lwd=2, bty="n", cex=1.5)

palette(viridis(75))
plot(cEigVectors[,15], mymerge[,niceTraitnames[myI]]/diag(sigma11), bty="n", xlab = "Conditional PC 15", ylab = "Difference btw breeding value and expectation", col=as.factor(mymerge$Latitude), lwd=2, main=niceTraitnames[myI])
legend('bottomleft', as.character(round(c(quantile(mymerge$Latitude, probs=seq(0,1,1/3), na.rm=T)))), pch=1, bty='n', col=viridis(75, direction=-1)[round(c(quantile(mymerge$Latitude, probs=seq(0,1,1/3), na.rm=T)))], pt.lwd=2)

palette(viridis(8))
plot(cEigVectors[,15], mymerge[,niceTraitnames[myI]]/diag(sigma11), bty="n", xlab = "Conditional PC 15", ylab = "Difference btw breeding value and expectation", col=as.factor(mymerge$Country), lwd=2)
legend("bottomleft", levels(as.factor(mymerge$Country)), col = viridis(8), pch=1, pt.lwd=2, bty="n")

myI = 16
palette(viridis(3))
plot(cEigVectors[,4], mymerge[,niceTraitnames[myI]]/diag(sigma11), bty="n", xlab = "Conditional PC 15", ylab = "Difference btw breeding value and expectation", col=as.factor(mymerge$Type), lwd=2, main=niceTraitnames[myI])
legend("topleft", levels(as.factor(mymerge$Type)), col = viridis(3)[1:2], pch=1, pt.lwd=2, bty="n", cex=1.5)

summary(mymerge$CobMass)

palette(viridis(8))
plot(cEigVectors[,4], mymerge[,niceTraitnames[myI]]/diag(sigma11), bty="n", xlab = "Conditional PC 15", ylab = "Difference btw breeding value and expectation", col=as.factor(mymerge$Country), lwd=2, main=niceTraitnames[myI])
legend("topleft", levels(as.factor(mymerge$Country)), col = viridis(8), pch=1, pt.lwd=2, bty="n")

```

```{r}
#lat long plot
library(maps)
eurodat = read.table('data/eurolandraceinfo.csv', sep=',', head=T, stringsAsFactors=F)

map("world", xlim=c(1.5*min(eurodat$Longitude),1.2*max(eurodat$Longitude)), ylim=c(0.85*min(eurodat$Latitude),1.05*max(eurodat$Latitude)), fill=T, col="azure")
points(eurodat$Longitude, eurodat$Latitude, col = magma(6)[4], lwd=4)

palette(magma(3))
par(xpd=TRUE)
map("world", xlim=c(1.5*min(eurodat$Longitude),1.2*max(eurodat$Longitude)), ylim=c(0.9*min(eurodat$Latitude),1.05*max(eurodat$Latitude)), fill=T, col="azure")
points(eurodat$Longitude, eurodat$Latitude, col = as.factor(eurodat$Type), lwd=4)
legend('bottomleft', levels(as.factor(eurodat$Type)), col = as.factor(eurodat$Type), pch=1, pt.lwd=4, bty="n")

```

## Inflation factor
-- essentially use the lower PCs to estimate Va's inflation due to whatever is going on.

```{r}
## pull out the Cm values
load('data/qxpc_euro_output.rda')
mycmprime = sapply(qxpceuroOut, function(x){x$cmprime}) #rows are PCs, columns are the traits

genicVa = sapply(qxpceuroOut, function(x){x$va})

## estimate Va with lower PCs and compare to genic Va
empVa = apply(mycmprime[(tailCutoff-50):tailCutoff,],2,var0) #get estimated Va for each trait

plot(log10(genicVa), log10(empVa), xlab = "log10 (genic Va)", ylab = "log10(estimated Va)", xlim = c(0,6), ylim=c(0,6), xaxt="n", yaxt = "n", bty = "l")
axis(1, at=log10(10^(0:6)), labels=10^(0:6))
axis(2, at=log10(10^(0:6)), labels=10^(0:6), las=2)

abline(a=0,b=1, lwd=2, col = viridis(3)[1])
abline(a=1/3,b=1, col = viridis(3)[2], lwd=2)
abline(a=1/2,b=1, col = viridis(3)[3], lwd=2)
legend('topleft', c('y = x','y = x + 1/3', 'y = x + 1/2'), lwd=2, col = viridis(3), bty="n")


##so estimated Va is generally higher than genic Va

plot(genicVa, empVa, xlab = "genic Va", ylab = "estimated Va",xaxt="n", yaxt = "n", bty = "l", xlim = c(0,8000), ylim=c(0,8000))
axis(1)
axis(2,las=2)
abline(a=0,b=1, lwd=2, col = viridis(3)[1])
abline(a=0,b=2, col = viridis(3)[2], lwd=2)
abline(a=0,b=3, col = viridis(3)[3], lwd=2)
legend('bottomright', c('y = x','y = 2x', 'y = 3x'), lwd=2, col = viridis(3), bty="n")

plot(genicVa, empVa, xlab = "genic Va", ylab = "estimated Va",xaxt="n", yaxt = "n", bty = "l", xlim = c(0,200), ylim=c(0,200))
axis(1)
axis(2,las=2)
abline(a=0,b=1, lwd=2, col = viridis(3)[1])
abline(a=0,b=2, col = viridis(3)[2], lwd=2)
abline(a=0,b=3, col = viridis(3)[3], lwd=2)
legend('bottomright', c('y = x','y = 2x', 'y = 3x'), lwd=2, col = viridis(3), bty="n")



##does a more stringent pvalue cutoff solve this problem
mycmprimec = sapply(qxpceuroOutcutoff, function(x){x$cmprime}) #rows are PCs, columns are the traits
genicVac = sapply(qxpceuroOutcutoff, function(x){x$va})

## estimate Va with lower PCs and compare to genic Va
empVac = apply(mycmprimec[(tailCutoff-50):tailCutoff,],2,var0) #get estimated Va for each trait

plot(log10(genicVac), log10(empVac), xlab = "log10 (genic Va)", ylab = "log10(estimated Va)", xlim = c(0,6), ylim=c(0,6), xaxt="n", yaxt = "n", bty = "l")
axis(1, at=log10(10^(0:6)), labels=10^(0:6))
axis(2, at=log10(10^(0:6)), labels=10^(0:6), las=2)

abline(a=0,b=1, lwd=2, col = viridis(3)[1])
abline(a=1/3,b=1, col = viridis(3)[2], lwd=2)
abline(a=1/2,b=1, col = viridis(3)[3], lwd=2)
legend('topleft', c('y = x','y = x + 1/3', 'y = x + 1/2'), lwd=2, col = viridis(3), bty="n")


```








Running the original version of the test.
```{r no-conditional-test}
#
load('data/euro-only-eigen.rda')

### function for running Qpc on European polygenic scores without the conditional test
Qxpceuro_nocond <- function(myI, myM = 906){
#read in data
gwasHits = read.table(paste('data/263-gwas-results/ldfiltered.assoc.',myI,sep=""), stringsAsFactors=F) #gwas results
names(gwasHits) = c('x','y',strsplit('chr     rs      ps      n_miss  allele1 allele0 af      beta    se      l_remle l_mle   p_wald  p_lrt   p_score', split=' +')[[1]])
gwasHits$locus =  sapply(gwasHits$rs, function(x){paste('s',gsub(":","_",x),sep="")})
sigGenos = read.table(paste('data/263-gwas-results/sigSnpsEuro.',myI, sep=""), header=T, stringsAsFactors=F) #genotypes of gwas snps in the European landraces

#combine table of GWAS results with genotypes in the GWAS set
combInfo = dplyr::left_join(sigGenos, gwasHits, by = 'locus')
combInfo$mybetas = ifelse(combInfo$allele1 == combInfo$ALT, combInfo$beta, -combInfo$beta)
myBetas = as.matrix(combInfo$mybetas)
myG = t(as.matrix(sigGenos[,6:(myM+5)]))

#center genotype matrix 
m = nrow(myG)
myT = matrix(data = -1/m, nrow = m - 1, ncol = m)
diag(myT) = (m - 1)/m
myGcent = myT %*% myG

#calculate breeding values
allZ = myGcent %*% myBetas

#project breeding values onto PCs and standardize by eigenvalue
myVa = calcVa(colMeans(myG)/2, myBetas)
myBm = t(allZ) %*% euroOnlyeigen$vectors

#do Qpc
myCmprime = sapply(1:(myM-1), function(x){t(myBm[,x]/sqrt(cEigValues[x]))})
myQm = sapply(1:pcmax, function(n){
    var0(myCmprime[n])/var0(myCmprime[(tailCutoff-50):tailCutoff])
  })
myPsprime = sapply(1:pcmax, function(x){pf(myQm[x], 1, 50, lower.tail=F)})

outList = list(cmprime = myCmprime, pprime = myPsprime, n.sites = nrow(combInfo))
}

### Run original Qpc on all polygenic scores
ncEuroOut = lapply(1:22,Qxpceuro_nocond)
save(ncEuroOut, file="data/qxpc-euro-nc.rda")

### look at the output
ncpvals = sapply(ncEuroOut, function(x) {x$pprime}) #matrix, rows are pvals, columns are traits
ncqvals = get_q_values(ncpvals)

#mysig = apply(allqvals, 2, function(x){ cut(x, c(0,0.001,0.01,0.05,0.1,1), labels=F)})
mysig2 =  cut((1:1000/1000), c(0,0.001,0.01,0.05,0.1,1)) #for legend
par(mar=c(8,14,2,2), xpd=TRUE, mfrow=c(1,1))
mycol = c(viridis(6, direction=1)[1:4], "white")
image(ncqvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
#axis(1, at = c(0:5/5), labels = (0:5)*10)
axis(1, at = c(0,0.2,0.4,0.6,0.8,1), labels=round(c(0,0.2,0.4,0.6,0.8,1)*nrow(ncqvals)))
axis(2, at=(0:21)/21, labels = niceTraitnames, las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)






````

