---
title: "Qpc-ames.Rmd"
author: "em"
date: "July 9, 2018"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(viridis)
library(qpctools)
library(qvalue)
setwd('~/Documents/qpc-maize/') #sorry world

```

This notebook has code for running a form of Q<sub>xpc</sub> analysis using the Major-Goodman GWAS panel and detecting selection in the Ames panel.

First, load in the kinship matrix and make a conditional matrix
```{r kmatrix}
load('data/ames.281E.K.rda')

sigma11 = as.matrix(myF[1:2704,1:2704])
sigma12 = as.matrix(myF[1:2704,2705:2984])
sigma21 = as.matrix(myF[2705:2984,1:2704])
sigma22 = as.matrix(myF[2705:2984,2705:2984]) #we are dropping the last row
sigma.cond = sigma11 - sigma12 %*% solve(sigma22) %*% sigma21 
```


```{r make-eigens, eval=F}
### This code makes a bunch of secondary sets of eigenvalues and saves them.
regEig = eigen(myF)
#write.table(regEig$vectors,file='data/ames.reg.eigenvectors', quote=F, row.names=F, col.names=F)
#write.table(regEig$values,file='data/ames.reg.eigenvalues', quote=F, row.names=F, col.names=F)
save(regEig, file = 'data/ames.281E.regeig.rda')

condEig = eigen(sigma.cond)
save(condEig, file = 'data/ames.281E.condeig.rda')

load('data/amesOnly.E.K.rda')
amesEig = eigen(amesOnly)
save(amesEig, file='data/amesOnly.eig.rda')

```

```{r read-eigens}
##read in the eigen data for the conditional amtrix
load('data/ames.281E.condeig.rda')
cEigVectors = condEig$vectors
cEigValues = condEig$values

#read in trait names
traitNames = read.table('data/blup.names', stringsAsFactors=F)$V1

```

```{r pccutoffs}

#get cutoffs for pcs based on % variation explained
varexp = cEigValues/sum(cEigValues)
sumexp = sapply(1:length(varexp), function(x){sum(varexp[1:x])})

#get cutoff with scree plot
library(nFactors)
myscree = nScree(cEigValues)

#get cutoffs for how many pcs to look at
pcmax = which(sumexp > 0.3)[1]

plot(cEigValues, bty="n", xlab = "PCs of conditional matrix", ylab = "Eigenvalues")
abline(v = pcmax, col = viridis(6)[3], lwd=2)
abline(v = myscree$Components[1], col = viridis(6)[4], lwd=2)


#remove the last end of PCs 
tailCutoff = round(.9*length(cEigValues))

```


# Run Qpc on the Ames Panel

The function is in qpctools/R/QpcAmes.R
```{bash}
cat qpctools/R/QpcAmes.R
```


```{r runqxpc, eval=F}
qpcamesOut = lapply(traitNames,Qpcames)
save(qpcamesOut, file='data/qpc-ames-output.rda')

```

# Look at results
```{r heatmap-figure}
load('data/qpc-ames-output.rda')

niceTraitnames = sapply(traitNames, function(x){
  strsplit(x, '0607')[[1]][1]
})

pcpvals = sapply(qpcamesOut, function(x) {x$pprime}) #matrix, rows are pvals, columns are traits
allqvals = get_q_values(pcpvals)

## TODO FIX AXES
layout(matrix(1, nrow=1, ncol=1))
mysig2 =  cut((1:1000/1000), c(0,0.001,0.01,0.05,0.1,1)) #for legend
par(mar=c(8,14,2,2), xpd=TRUE, mfrow=c(1,1))
mycol = c(viridis(6, direction=1)[1:4], "white")
image(allqvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
#axis(1, at = c(0:5/5), labels = (0:5)*10)
axis(1, at = c(0,0.2,0.4,0.6,0.8,1), labels=round(c(0,0.2,0.4,0.6,0.8,1)*nrow(allqvals)))
axis(2, at=(0:21)/21, labels = niceTraitnames, las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

#mysig = apply(allqvals, 2, function(x){ cut(x, c(0,0.001,0.01,0.05,0.1,1), labels=F)})
image(pcpvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
#axis(1, at = c(0:5/5), labels = (0:5)*10)
axis(1, at = c(0,0.2,0.4,0.6,0.8,1), labels=round(c(0,0.2,0.4,0.6,0.8,1)*nrow(allqvals)))
axis(2, at=(0:21)/21, labels = niceTraitnames, las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

```


# Running the original version of the test.

The function is in qpctools/R/QpcAmes-nocond.R
```{bash}
cat qpctools/R/QpcAmes-nocond.R
```


```{r no-conditional-test, eval=F}
#
load('data/amesOnly.eig.rda')

ncamesOut = lapply(traitNames, Qpcames_nocond)
save(ncamesOut, file = 'data/qpc-ames-nc.rda')

```

```{r}

load('data/qpc-ames-nc.rda')

ncpvals = sapply(ncamesOut, function(x) {x$pprime}) #matrix, rows are pvals, columns are traits
ncqvals = get_q_values(ncpvals)

## TODO FIX AXES
#mysig = apply(allqvals, 2, function(x){ cut(x, c(0,0.001,0.01,0.05,0.1,1), labels=F)})
layout(matrix(1, nrow=1, ncol=1))
mysig2 =  cut((1:1000/1000), c(0,0.001,0.01,0.05,0.1,1)) #for legend
par(mar=c(8,14,2,2), xpd=TRUE, mfrow=c(1,1))
mycol = c(viridis(6, direction=1)[1:4], "white")
image(ncqvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at = c(0,0.2,0.4,0.6,0.8,1), labels=round(c(0,0.2,0.4,0.6,0.8,1)*nrow(allqvals)))
axis(2, at=(0:21)/21, labels = niceTraitnames, las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

image(ncpvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
#axis(1, at = c(0:5/5), labels = (0:5)*10)
axis(1, at = c(0,0.2,0.4,0.6,0.8,1), labels=round(c(0,0.2,0.4,0.6,0.8,1)*nrow(allqvals)))
axis(2, at=(0:21)/21, labels = niceTraitnames, las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)


save(qpcamesOut, allqvals, niceTraitnames, ncqvals, ncamesOut, file="data/ames_qpc_data.rda")
````

