---
title: "Simulations_qpc.Rmd"
author: "em"
date: "February 13, 2017"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(viridis)
library(qpctools)
library(qvalue)
library(MASS)
library('LaCroixColoR')
setwd('~/Documents/qpc-maize/') #sorry world

myBound <- function(x){
if (x>1) {x=1}
if (x<0) {x=0}
  return(x)
}

getPopGenos <- function(p, pops, popSize){ #turns allele frequencies into individual genotypes, p is row, pops is the matrix of frequencies, popSize is the # of individuals per pop
  indFreqs = t(matrix(rep(pops[p,],popSize),nrow=nloci,ncol=popSize))
  popGenos = apply(indFreqs,c(1,2),function(x){sum(sample(c(1,0),2,replace=TRUE, prob = c(x,1-x)))})
  return(popGenos)
}
```

This notebook has code for running simulations for Q<sub>pc</sub> analysis in a GWAS panel of 240 maize lines and Q<sub>xpc</sub> on Ames and European lines.


###Q<sub>pc</sub> simulations:

```{r qpc_sims}
#read in kinship data
myK = read.table('data/All_240E.nomaf.nomissing.K')
myKnames = read.table('data/240.names', stringsAsFactors = F)$V1 #get individual names
row.names(myK) = myKnames[1:dim(myK)[1]]

#eigendecomposition of the kinship matrix
eigF = eigen(myK)
myU = eigF$vectors
myLambdas = eigF$values

#get test cutoffs
varexp = myLambdas/sum(myLambdas)
sumexp = sapply(1:length(varexp), function(x){sum(varexp[1:x])})
pcmax = which(sumexp > 0.3)[1]
```

```{r qpcsims, eval=F}


#simulate traits with neutrality and varying amounts of Va. Run this once and save output to save time.
neutQpc <- function(myK,myU,myLambdas, myseed=1,myVa=1, myVe=0){
set.seed(myseed)
myZneutral = mvrnorm(n=1, rep(0,dim(myK)[1]), myVa*myK + myVe*diag(dim(myK)[1])) #neutral traits
myQpcneutral = calcQpc(myZneutral, myU, myLambdas, tailCutoff=1, vapcs = 119)
return(myQpcneutral)
}
neutQpcVe0 = lapply(1:200, function(x){neutQpc(myK, myU, myLambdas, myseed=x, myVa=1, myVe=0)})#no Ve
neutQpcVe.1 = lapply(1:200, function(x){neutQpc(myK, myU, myLambdas, myseed=x,myVa=1, myVe=.1)}) #Ve = Va/10
neutQpcVe.5 = lapply(1:200, function(x){neutQpc(myK, myU, myLambdas, myseed=x,myVa=1, myVe=.5)}) #Ve = Va/2

save(neutQpcVe0,neutQpcVe.1,neutQpcVe.5, pcmax,file="data/simFiles/Qpc200.rda")
```

We can measure false positives using the variance of C values (which we expect to be Va under neutrality, which has been set as 1 here)
```{r qpc-sims-cont}
load('data/simFiles/Qpc200.rda')
mycol = viridis(6)


#no Ve
allCms = sapply(neutQpcVe0, function(x) {x$cm}) #matrix where columns are the Cm values for a given simulation (from PC1 to PC 239)
allCmVars = apply(allCms, 1, var0) #get variance across all 200 sims for each pc
plot(allCmVars[1:238], bty="n", col = mycol[1], lwd=2, xlab = 'PC', ylab = "var(Cm) with no Ve", xlim = c(0,240))

allPs = sapply(neutQpcVe0, function(x) {x$pvals}) #matrix where columns are the Cm values for a given simulation (from PC1 to PC 239)

count_sig = sapply(1:pcmax, function(x){sum(allPs[x,]<=0.05)})
```

Without any V<sub>e</sub>, the variance of C<sub>m<sub> is centered around 1, as expected under neutrality. What happpens when we add V<sub>e</sub>?

```{r}
#adding Ve
allCmsVe.1 = sapply(neutQpcVe.1, function(x) {x$cm})
allCmVarsVe.1 = apply(allCmsVe.1, 1, var0) #get variance across all 200 sims for each pc
allCmsVe.5 = sapply(neutQpcVe.5, function(x) {x$cm})
allCmVarsVe.5 = apply(allCmsVe.5, 1, var0) #get variance across all 200 sims for each pc

plot(allCmVarsVe.5[1:238], bty="n", col = mycol[1], lwd=2, xlab = 'PC', ylab = "var(Cm)", xlim = c(0, 238), ylim=c(0,3.5))
points(allCmVarsVe.1[1:238], col = mycol[3], lwd=2)
points(allCmVars[1:238], col = mycol[5], lwd=2)
legend('topleft', c('Ve=0','Ve=Va/10','Ve=Va/2'), col = mycol[c(5,3,1)], pch=1, bty="n", pt.lwd=2)

```

Adding V<sub>e</sub> increases the variance of Cm at increasingly large scales. This will make our test conservative. We can see how conservative but calculating the average P value across tests. 

```{r}
#testing for selection on early PCs
allPs = sapply(neutQpcVe0, function(x){x$pvals})
meanPs = apply(allPs, 1, mean)
allPsVe.1 = sapply(neutQpcVe.1, function(x){x$pvals})
meanPsVe.1 = apply(allPsVe.1, 1, mean)
allPsVe.5 = sapply(neutQpcVe.5, function(x){x$pvals})
meanPsVe.5 = apply(allPsVe.5, 1, mean)

plot(-log10(meanPs), col = mycol[1], bty="n", xlab = "PC", ylab = "-log10(p)", ylim=c(0.1,0.5), lwd=2)
points(-log10(meanPsVe.1), col=mycol[3], lwd=2)
points(-log10(meanPsVe.5), col=mycol[5], lwd=2)
legend('topleft', c('Ve=0','Ve=Va/10','Ve=Va/2'), col = mycol[c(1,3,5)], pch=1, bty="n", pt.lwd=2)
```

Doing simulations where we incorporate actual loci

```{r, eval=F}


## simulate allele frequencies in each of 240 pops. nloci = 10050, 50 of these are used to make the trait, the rest to estimate K
nloci = 100500
mysims = sapply(1:200, function(i){
set.seed(i)
ancPop = runif(nloci, min=0, max=1) #ancestral allele frequenices
presentPops1 = sapply(ancPop, function(x){mvrnorm(n=1, mu = rep(x,239), x*(1-x)*myK)})
presentPops = apply(presentPops1, c(1,2), myBound) #deal with numbers greater or less than 0 (the outer bounds are sticky)
popGenos = sapply(1:239, function(x) getPopGenos(x, presentPops, 1)) # matrix where each row is a SNP and each column is an individual
myG = t(popGenos[501:100500,])/2#snps used to make K matrix
myKsim = make_k_E(myG) #make a new pop matrix based on the simulated snps
myEigsim = eigen(myKsim)
beetas = matrix(c(rnorm(50), rep(0, 10000)), ncol=1, nrow=nloci) 
popPhenos = apply(popGenos, 2, function(x){x %*% beetas})
myQpcneutral = calcQpc(popPhenos, myEigsim$vectors, myEigsim$values, tailCutoff=1, vapcs= 119)
return(list(phenos = popPhenos, kSim = myKsim, myCms = myQpcneutral$cm))
})

save(mysims, file = "data/simFiles/qpc-loci-sims.rda")


```


Look at the simulations



```{r}
mycol = lacroix_palette('Mango')
load('data/simFiles/qpc_loci_sims.rda')
load('data/simFiles/qpc_loci_sims500.rda')
#mysims is a matrix, each column is a sim. first row is phenotypes (in 1 item list), second row is a one item list with the kinship matrix in it, third row is a list of the cm values. 
#what do CM values look like across PCs
myCms = sapply(1:200, function(x){mysims[3,x][[1]]}) #rows are PCs, cols are simulations
myCmMeans = sapply(1:238, function(x){mean(myCms[x,]^2)})

#generate the Cm values for the 500 snp sims
myQpc500 = sapply(1:200, function(i){
  myphenos = mysims500[[i]]
  kSim = mysims[,i]$kSim
  myEigsim = eigen(kSim)
  myQpcneutral = calcQpc(myphenos, myEigsim$vectors, myEigsim$values)
  return(myQpcneutral)
})

myCms500 = sapply(1:200, function(x){myQpc500[1,x][[1]]}) #rows are PCs, cols are simulations
myCmMeans500 = sapply(1:238, function(x){mean(myCms500[x,]^2)})


#what is Va going to be?? based on our expectations from neutral sites
myVas = sapply(1:200, function(x){calcVa(runif(50, min=0, max=1), rnorm(50))})
hist(myVas)

myVas500 = sapply(1:200, function(x){calcVa(runif(500, min=0, max=1), rnorm(500))})
hist(myVas500)

## so Cm~N(0,2Va), so cm^2 ~ 2Va

#for the 50 snps sims
plot(-100,-100, xlim = c(0,240), ylim = c(0,max(myCms)^2), bty="n", xlab = "PC", ylab = "Cm^2")
test = sapply(1:200, function(x){points(1:238,myCms[,x]^2, col = mycol[3])})
abline(v = 0.9*239, lwd=2)
points(1:238,myCmMeans, col = mycol[6], pch=16)

myl = lm(myCmMeans ~ c(1:238))
summary(myl)

noisefit = 15/myLambdas

plot(1:238, myCmMeans, bty="n", xlab = "PC", ylab = "CM^2 means")
abline(myl, col = mycol[2], lwd=3)
abline(h = mean(myVas)*2, lwd=3, col = mycol[4])
points(noisefit, type='l', lwd=2, col = mycol[3])

legend('topleft', c('Va', 'linear model', 'noise'), lwd=2, col = mycol[c(4,2,3)], bty="n")


## for the 500 snps sims
myl500 = lm(myCmMeans500 ~ c(1:238))
summary(myl500)


plot(1:238, myCmMeans500, bty="n", xlab = "PC", ylab = "CM^2 means")
abline(myl500, col = mycol[2], lwd=3)
abline(h = mean(myVas500)*2, lwd=3, col = mycol[4])
#points(noisefit, type='l', lwd=2, col = mycol[3])

#legend('topleft', c('Va', 'linear model', 'noise'), lwd=2, col = mycol[c(4,2,3)], bty="n")
legend('topleft', c('Va', 'linear model'), lwd=2, col = mycol[c(4,2)], bty="n")


#how often are we detecting selection using our standard parameters
myQpc = sapply(1:200, function(x){calcQpc(mysims[1,x]$phenos, eigen(mysims[2,x]$kSim)$vectors, eigen(mysims[2,x]$kSim)$values)})

myps = sapply(1:200, function(x){myQpc[,x]$pvals})
plot(-100,-100, xlim = c(0,40), ylim=c(0, -log10(min(unlist(myps)))*1.1), bty="n", xlab = "PC", ylab = "-log10 p")
test = sapply(1:200, function(x){points(-log10(myps[[x]]))})

#how many are below 0.05
myps0.05 = sapply(1:40, function(x){
  pcpvals = unlist(lapply(myps, function(y){y[x]}))
  return(sum(pcpvals<0.05, na.rm=T))
})
plot(myps0.05/200, bty="n", xlab = "PC", ylab = "proportion where p<0.05", ylim = c(0,0.25))
abline(h=0.05, col = mycol[3], lwd=2)



#how often are we detecting selection using our standard parameters for 500 snps

myps500 = sapply(1:200, function(x){myQpc500[,x]$pvals})
plot(-100,-100, xlim = c(0,40), ylim=c(0, -log10(min(unlist(myps500)))*1.1), bty="n", xlab = "PC", ylab = "-log10 p")
test = sapply(1:200, function(x){points(-log10(myps500[[x]]))})

#how many are below 0.05
myps0.05 = sapply(1:40, function(x){
  pcpvals = unlist(lapply(myps500, function(y){y[x]}))
  return(sum(pcpvals<0.05, na.rm=T))
})
plot(myps0.05/200, bty="n", xlab = "PC", ylab = "proportion where p<0.05", ylim = c(0,0.25))
abline(h=0.05, col = mycol[3], lwd=2)


##to do -- compare Va estimatates across all sites...

```
































